Pathogen Spillover - dissertation chapter

### 2018-01-11 pathogen-spillover-diss-ch.Rmd
Don't evaluate some models for document output.

[Conceptual Path Model](figures/conceptual-path-model-updated.png)

@@

```{r load data, echo=F, message=F}
library(knitr)
load("climate_path_model_data.RData")
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
library(plyr); library(dplyr, verbose = F)
library(lme4, verbose = F)
library(lmerTest, verbose = F)

@@

## make consolidated data frame only of variables in the path model
oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct.log, tot_lfct.log.scl, hrs_abv25ds, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log, umca_density_ha.log.scl, H.2005, shannons2005.scl, twi15m.log, twi15m.log.scl)

oakplots.model.data <- filter(oakplots.model.data, is.na(H.2005)==F)
write.csv(oakplots.model.data, "analysis/data/oakplots-pathmodel-data.csv",
          row.names = F)
```

@@

```{r hours >25 path model no umca density unscaled, echo=F}
model_list2 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.model.data, sample_year == 2006)),

@@

```{r hours >25 path model no umca density scaled, echo=F}
model_list2.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),

@@

```{r hours >25 path model all oak plots unscaled, echo=F}

model_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),

@@

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ 
                                   # umca_density_ha.log + twi15m.log +
                                   twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      hrs_abv25 <- lmer(hrs_abv25ds ~ 
                              # umca_density_ha.log + 
                              twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),

```{r hours >25 path model all oak plots scaled}
model_list1.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),

@@

In my assessment of model residuals I found no substantial violations of the assumption of homoscedacity for individual models, and no relationships between the model residuals and the predictor variable(s). The patterning in the residual plots for some of the models is likely in part due to observations of count variables where the value is zero. Since the dataset in this analysis is all plots with one or more susceptible oak species there are some plots that have zero bay laurel and zero symptomatic bay laurel leaves.

@@

**Residual plots of Shannon's Diversity Index model**
```{r shannons residuals plot, echo=F}
plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
plot(shannons$model$twi15m.log.scl, resid(shannons),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Diversity model residuals vs. wetness index predictor")
```

**Residual plots of bay laurel density model**
```{r bay laurel density residuals plot, echo=F}
plot(umca_density, main = "LMM of bay laurel density residuals")
plot(umca_density@frame$twi15m.log.scl, resid(umca_density),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Bay laurel density model residuals vs. wetness index predictor")
```

@@

plot(slc@frame$umca_density_ha.log.scl, resid(slc),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "SLC model residuals vs. bay laurel density predictor")
plot(slc@frame$hrs_abv25ds.scl, resid(slc),
     xlab = "# dry-season hours >25 C",
     ylab = "model residuals",
     main = "SLC model residuals vs. dry-season hours >25 C predictor")
plot(slc@frame$avg_hrs1422_wet_t_t1.log.scl,
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "SLC model residuals vs. wet-hours predictor")
plot(slc@frame$shannons2005.scl, resid(slc),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "SLC model residuals vs. Shannon's diversity predictor")
# plot(slc@frame$twi15m.log.scl, resid(slc),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. wetness index predictor")
```
```{r pairs plot of slc predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, umca_density_ha.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl, twi15m.log.scl)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for symptomatic leaf count models")
```

**Residual plots of average number of wet-hours model**
```{r wet-hours residuals plot, echo=F}
plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
# plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "Wet-hours model residuals vs. bay laurel density predictor")
plot(wet_hours_1422@frame$twi15m.log.scl, resid(wet_hours_1422),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "Wet-hours model residuals vs. bay laurel density predictor")
```

**Residual plots of dry-season temperature >25C model**
```{r dry-season hours residuals plot, echo=F}
plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
plot(hrs_abv25@frame$twi15m.log.scl, resid(hrs_abv25),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Dry-season hours model residuals vs. wetness index predictor")
# plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Dry-season hours model residuals vs. wetness index predictor")
```

**Residual plots of oak infection model**
```{r oak infection residuals plot, echo=F}
plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
plot(oak_infection@frame$tot_lfct.log.scl, resid(oak_infection),
     xlab = "symptomatic leaf count",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log.scl, resid(oak_infection),
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. wet-hours predictor")
plot(oak_infection@frame$shannons2005.scl, resid(oak_infection),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. diversity predictor")
```
```{r pairs plot of oak infection predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, tot_lfct.log.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for oak infection model")
```

@@

```{r fit and compare all oak plots, echo=F, message=F, warning=F}
raw_variables_model <- sem.fit(model_list1, oakplots.model.data, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
        conditional = T, .progressBar = F)
scaled_variables_model <- sem.fit(model_list1.scl, oakplots.model.data,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
          conditional = T, .progressBar = F)

scaled_variables_modela <- sem.fit(model_list1.scla, oakplots.model.data,
          # corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
          conditional = F, .progressBar = F)

@@

```{r tables of d-sep test results from sem.fit, echo=F}
kable(raw_variables_model, digits = 4,
      caption = "Results of the d-sep test using raw variables.")
kable(scaled_variables_model,
      caption = "Results of the d-sep test using standardized variables.")

@@

```{r get path coefficients, echo=F}
raw_coefficients <- sem.coefs(model_list1, oakplots.model.data,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_coefficients <- sem.coefs(model_list1.scl, oakplots.model.data,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))

@@

@@

(sstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds), notch = T) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Heat Exposure (dry-season Hours >25 C)")
)

(rstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs1422_wet), notch = T) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Warm & Wet Conditions (wet-hours 14-22 C)")
)

@@

```{r plot rainfall figure, echo=F}
library(ggplot2)
# summary(rainy_season_ppt)
# rainy_season_ppt$year <- as.factor(rainy_season_ppt$year)
rainfall_figure <- ggplot(filter(rainy_season_ppt, year > 2004), 
                          aes(year, rain2d_total))
rainfall_figure +
      geom_boxplot(aes(group = year), notch = T) +
      geom_hline(aes(yintercept = mean(rain2d_total)),
                 colour = "red") +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Total Rainfall (cm)") +
      ggtitle("Rainfall Estimated at Plots from Rain Gauges (November - May)")
```

### Session Information
```{r print session info, echo=FALSE}
print(sessionInfo())
```

>>>>

![Conceptual Path Model](figures/temperature-path-model-updated.png)

@@

```{r set options, echo=F}
# knitr::opts_chunk$set(fig.width = 7.5, fig.height = 7)
knitr::opts_chunk$set(dpi = 300)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```

```{r load data, echo=F, message=F}
library(knitr)
load("climate_path_model_data.RData")
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
## Avoided this error in 1.1.2 by removing observations with NA values, which gives the same results as before, so maybe it wasn't properly dropping missing values when extracting info from the models.
library(plyr); library(dplyr, verbose = F)
library(lme4, verbose = F); library(nlme, verbose = F)
# library(lmerTest, verbose = F)

@@

## make consolidated data frame only of variables in the path model
oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
                              )

oakplots.model.data <- filter(oakplots.model.data, avg_hrs1422_wet_t_t1 !="NA", H.2005 != "NA", tot_lfct != "NA")
write.csv(oakplots.model.data, "analysis/data/oakplots-pathmodel-data.csv",
          row.names = F)

oakplots.model.data.sub <- oakplots.model.data %>% 
      filter(plotid!="hood03", plotid!="hood05", plotid!="jlsp04", plotid!="johns02", plotid!="sugar17")

write.csv(oakplots.model.data.sub, "pathogen_spillover_data.csv", row.names = F)
```

@@

```{r hours >25 path model no umca density unscaled, echo=F, eval=F}
model_list2 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.model.data, sample_year == 2006)),

@@

```{r hours >25 path model no umca density scaled, echo=F, eval=F}
model_list2.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),

@@

```{r hours >25 path model all oak plots unscaled, echo=F}
# slc_nlme <- lme(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005, random = ~1|plotid, data = oakplots.model.data)

model_list1 <- list(
      shannons <- lm(
            H.2005 ~ twi15m, 
            data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density <- lmer(
            umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

@@

      wet_hours_1422 <- lmer(
            avg_hrs1422_wet_t_t1.log ~ twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

      hrs_abv25 <- lmer(
            hrs_abv25ds ~ twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

      oak_infection <- glmer(
            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), 
            data = oakplots.model.data)
)

# refit model with missing path from diversity to umca density
model_list1_refit <- list(
      shannons_refit <- lm(
            H.2005 ~ twi15m, 
            data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density_refit <- lmer(
            umca_density_ha.log ~ H.2005 + twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

      slc_refit <- lmer(
            tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
                  data = oakplots.model.data),

      wet_hours_1422_refit <- lmer(
            avg_hrs1422_wet_t_t1.log ~ twi15m + (1|plotid) + (1|sample_year), 
                             data = oakplots.model.data),

      hrs_abv25_refit <- lmer(
            hrs_abv25ds ~ twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

      oak_infection_refit <- glmer(
            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
            family = binomial(link = "logit"), 
            data = oakplots.model.data)
)
```

```{r path-model-no-lide-plots, echo=F}
model_list1.noLIDE <- list(
      shannons <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data.sub, sample_year == 2006)),

      umca_density <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),

      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ 
                                   # umca_density_ha.log + twi15m.log +
                                   twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),

      hrs_abv25 <- lmer(hrs_abv25ds ~ 
                              # umca_density_ha.log + 
                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data.sub)
)

```{r hours >25 path model all oak plots scaled, eval = F}
model_list1.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),

@@

In my assessment of model residuals I found no substantial violations of the assumption of homoscedacity for individual models (although there was some), and no relationships between the model residuals and the predictor variable(s). The patterning in the residual plots for some of the models is likely in part due to observations of count variables where the value is zero. Since the dataset in this analysis is all plots with one or more susceptible oak species there are some plots that have zero bay laurel and zero symptomatic bay laurel leaves.

@@

**Residual plots of Shannon's Diversity Index model**
```{r shannons residuals plot, echo=F}
plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
plot(shannons$model$twi15m, resid(shannons),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Diversity model residuals vs. wetness index predictor")
```

**Residual plots of bay laurel density model**
```{r bay laurel density residuals plot, echo=F}
plot(umca_density, main = "LMM of bay laurel density residuals")
plot(umca_density@frame$twi15m, resid(umca_density),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Bay laurel density model residuals vs. wetness index predictor")
```

@@

plot(slc@frame$umca_density_ha.log, resid(slc),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "SLC model residuals vs. bay laurel density predictor")
plot(slc@frame$hrs_abv25ds, resid(slc),
     xlab = "# dry-season hours >25 C",
     ylab = "model residuals",
     main = "SLC model residuals vs. dry-season hours >25 C predictor")
plot(slc@frame$avg_hrs1422_wet_t_t1.log,
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "SLC model residuals vs. wet-hours predictor")
plot(slc@frame$H.2005, resid(slc),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "SLC model residuals vs. Shannon's diversity predictor")
# plot(slc@frame$twi15m.log.scl, resid(slc),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. wetness index predictor")
```
```{r pairs plot of slc predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, umca_density_ha.log, hrs_abv25ds, avg_hrs1422_wet_t_t1.log, H.2005, twi15m)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for symptomatic leaf count models")
```

**Residual plots of average number of wet-hours model**
```{r wet-hours residuals plot, echo=F}
plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
# plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "Wet-hours model residuals vs. bay laurel density predictor")
plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "Wet-hours model residuals vs. bay laurel density predictor")
```

**Residual plots of dry-season temperature >25C model**
```{r dry-season hours residuals plot, echo=F}
plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Dry-season hours model residuals vs. wetness index predictor")
# plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Dry-season hours model residuals vs. wetness index predictor")
```

**Residual plots of oak infection model**
```{r oak infection residuals plot, echo=F}
plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
     xlab = "symptomatic leaf count",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. wet-hours predictor")
plot(oak_infection@frame$H.2005, resid(oak_infection),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. diversity predictor")
```
```{r pairs plot of oak infection predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, tot_lfct.log, avg_hrs1422_wet_t_t1.log, H.2005)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for oak infection model")
```

@@

```{r fit and compare all oak plots, echo=F, message=F, warning=F}
raw_variables_model <- sem.fit(model_list1, oakplots.model.data, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
        conditional = T, .progressBar = F)

raw_vars_refit <- sem.fit(model_list1_refit, oakplots.model.data, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
        conditional = T, .progressBar = F)

## Model with all LIDE plots excluded
# sem.fit(model_list1.noLIDE, oakplots.model.data.sub, 
#         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
#         conditional = T, .progressBar = F)

# scaled_variables_model <- sem.fit(model_list1.scl, oakplots.model.data,
#           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
#           conditional = T, .progressBar = F)
# scaled_variables_modela <- sem.fit(model_list1.scla, oakplots.model.data,
#           # corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
#           conditional = F, .progressBar = F)

@@

```{r tables of d-sep test results from sem.fit, echo=F}
kable(raw_variables_model, digits = 4,
      caption = "Results of the d-sep test using raw variables.")
# kable(scaled_variables_model,
#       caption = "Results of the d-sep test using standardized variables.")

@@

```{r get path coefficients, echo=F}
raw_coefficients <- sem.coefs(model_list1, oakplots.model.data,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_coefficients <- sem.coefs(model_list1, oakplots.model.data,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
          standardize = "scale")

raw_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
          standardize = "scale")

## Model with LIDE plots excluded
# sem.coefs(model_list1.noLIDE, oakplots.model.data.sub,
#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
# sem.coefs(model_list1.noLIDE, oakplots.model.data.sub,
#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
#           standardize = "scale")

@@

By calculating oak infection at the plot-level we aggregated any variation in susecptibility that may be present in different species or individuals.

@@

(sstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds), notch = T) + 
      geom_hline(aes(yintercept = mean(hrs_abv25ds)), colour = "red") +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Heat Exposure (dry-season Hours >25 C)") +
      theme(panel.grid = element_blank())
)

(rstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs1422_wet), notch = T) +
      geom_hline(aes(yintercept = mean(hrs1422_wet)), colour = "red") +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Warm & Wet Conditions (wet-hours 14-22 C)") +
      theme(panel.grid = element_blank())
)
```

```{r plot rainfall figure, echo=F}
library(ggplot2)
rainfall
# summary(rainy_season_ppt)
# rainy_season_ppt$year <- as.factor(rainy_season_ppt$year)
rainfall_figure <- ggplot(filter(rainy_season_ppt, year > 2004), 
                          aes(year, rain2d_total))
rainfall_figure +
      geom_boxplot(aes(group = year), notch = T) +
      geom_hline(aes(yintercept = mean(rain2d_total)), linetype = "dashed") +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Total Rainfall (mm)") +
      # ggtitle("Rainfall Estimated at Plots from Rain Gauges (November - May)") +
      theme(panel.grid = element_blank(),
            axis.title = element_text(size = 12),
            axis.text = element_text(size = 10))
# ggsave("figures/rainy-season-ppt.eps", width = 6.25, height = 4.5)

rainy_days_figure <- ggplot(filter(rainy_season_ppt, year > 2004), 
                          aes(year, rain_days_v))
rainy_days_figure +
      geom_boxplot(aes(group = year), notch = F) +
      geom_hline(aes(yintercept = mean(rain_days_v)),
                 colour = "red") +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of days w/precipitation") +
      ggtitle("Rainy Days at Plots (November - May)")
```

## Session Information
```{r print session info, echo=FALSE}
print(sessionInfo())
```

## Model a subset of the data

### Bay laurel density ~ Diveristy relationship

```{r}
ggplot(oakplots.model.data,
       aes(umca_density_ha, H.2005)) +
      geom_point() +
      geom_smooth(method = "loess") +
      facet_grid(sample_year~.) +
      theme_bw()

d1 <- select(oakplots.sub, plotid:inf_oak_ct, H.2005)
d1 <- mutate(d1, tot_oak = rowSums(cbind(inf_oak_ct, uninf_oak_ct)))
ggplot(filter(d1, sample_year == 2006), aes(log1p(tot_oak), H.2005)) +
      geom_point() +
      geom_smooth(method = "lm") +
      theme_bw()

@@

I want to examine the different matrices of the mixed effects models using a smaller datset that will make them easier to visualize.
```{r model subset data, eval=F}
small_data <- oakplots.model.data[1:103,]
xtabs(~sample_year + plotid, droplevels(small_data))
## shows a somewhat unbalanced data set

small_data_model_list <- list(
      shannons_small <- lm(H.2005 ~ twi15m, data = filter(small_data, sample_year == 2006)),# this model avoids pseudoreplication

      umca_density_small <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = small_data),

      slc_small <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = small_data),

      wet_hours_1422_small <- lmer(avg_hrs1422_wet_t_t1.log ~ 
                                   # umca_density_ha.log + twi15m.log +
                                   twi15m + (1|plotid) + (1|sample_year), data = small_data),

      hrs_abv25_small <- lmer(hrs_abv25ds ~ 
                              # umca_density_ha.log + 
                              twi15m + (1|plotid) + (1|sample_year), data = small_data),

      oak_infection_small <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = small_data)
)

summary(umca_density_small)
summary(slc_small)
summary(hrs_abv25_small)
anova(umca_density,update(umca_density, .~ 1+. -twi15m))

getME(umca_density_small, "X")# fixed-effects model matrix
getME(umca_density_small, "Z")# random-effects model matrix
getME(umca_density_small, "Zt")# transposed random-effects model matrix, diagonal values = 1
# image(getME(umca_density_small, "Z"))#
getME(umca_density_small, "Ztlist")# list of components of Zt separated by individual variance component
getME(umca_density_small, "mmList")# list of raw model matrices associated w/random effects terms
getME(umca_density_small, "u")# conditional mode of "spherical" random effects variable
getME(umca_density_small, "b")# conditional mode of random effects variable
getME(umca_density_small, "Gp")

image(getME(umca_density_small, "L"))# sparse Cholesky factor matrix of the penalized random-effects model
image(getME(umca_density_small, "Lambda"))# relative covariance factor of the random effects
image(getME(umca_density_small, "Lambdat"))# cross-product of random effects model matrix, diagonal values = random-effects parameter estimates

getME(umca_density_small, "A")# scaled sparse model matrix
getME(umca_density_small, "RX")# Cholesky factor for fixed-effects parameters
getME(umca_density_small, "RZX")#cross-term in the full Cholesky factor

getME(umca_density_small, "sigma")# residual standard error

getME(umca_density_small, "theta")# random-effects parameter estimates

getME(umca_density_small, "q")# number of coluns of RE model matrix Z
getME(umca_density_small, "l_i")# number of levels of grouping factors
getME(umca_density_small, "k")# number of RE terms
getME(umca_density_small, "m_i")# number of covariance parameters in each RE term
getME(umca_density_small, "devcomp")# list of items describing fitted model

# image(getME(oak_infection, "L"))
```
